Introduction to LLL\\``Cryptography''

### 2021-05-27 paper.tex
RSA Attack TODO: modify parameter

\begin{center}
    \begin{eqnarray}
         g(x_0) \equiv 0 \mod N^m, |x_0| \le X \\
         \lVert g(xX) \rVert < \frac{N}{\sqrt{n}} \label{eq:pol}
    \end{eqnarray}
\end{center}

Then $g(x_0) = 0$ holds over the integers.

@@

Suppose that we don't know the factorization of $N$ and we would like to know the message $m$.
However we know the padding and the length of $m < 2^{16}$.

@@

which is the known padding string.

@@

\begin{align*}
    f(x) = x^3 + 0x15d393c596142306bae0000x^2 + 0x1b53c5e184a49b39f9ad9eedbx \\
    + 0x9bb23cefd94fac9cc9cee9943
\end{align*}

@@

If we apply LLL to the lattice we get

\clearpage
We can use lattices to solves the problem:

\begin{itemize}
    \item If we have some polynomial which share the same root $x_0$ over $N^m$ we can represent each of them as a row from lattice,
        and every linear integer combination of rows between them will yield another polynomial having a root $x_0$.
    \item Using \textbf{LLL} on the lattice we can find a reduced lattice basis, such that the norm of the shortest vector will
        satisfy the inequality of \ref{eq:pol}. \textbf{LLL} will only use \textbf{linear integer operations} on the basis vectors,
        in this way we preserve the root $x_0$ and the bound $X$.
    \item Define the shortest vector in the reduced basis as $v = (v_0, v_1, \ldots, v_n)$. We can construct the polynomial
        \begin{center}
            \[
                g(x) = v_0 + \frac{\displaystyle v_1}{X}x + \frac{v_2}{X^2}x^2 + \cdots + \frac{v_n}{X^n}x^n
            \]
        \end{center}
\end{itemize}
Since $g(x)$ is on the lattice we know that

\begin{center}
    \begin{eqnarray}
        g(x_0) \equiv 0 \mod N^m \\
        \vert x_0| \le X \\
        \lVert g(xX) \rVert < \frac{N^m}{\sqrt{deg(g) + 1}}
    \end{eqnarray}
\end{center}

>>>>

\begin{center}
    \begin{eqnarray}
        g(x_0) \equiv 0 \mod N^m, |x_0| \le X \label{eq:bound} \\
         \lVert g(xX) \rVert < \frac{N}{\sqrt{n}} \label{eq:pol}
    \end{eqnarray}
\end{center}

\textit{Then $g(x_0) = 0$ holds over the integers.}

This theorem states that is possible to compute the root of $f(x) \mod N$ if we can find a polynomial $g(x)$ that share the same root but modulo $N^m$.
If \ref{eq:bound} and \ref{eq:pol} hold then we can simply compute the root of $g(x)$ over the integers to have the same root $x_0$
such that $f(x_0) \equiv 0 \mod N$.

Howgrave-Graham's idea is to find this polynomial $g$ by combining polynomials $g_{i,j}$ who also have $x_0$ as roots modulo $N^m$.

The \textbf{LLL} algorithm is fundamental because:

\begin{itemize}
    \item It only does \textbf{integer linear operations} on the basis vectors.
        In this way even if the basis is different it's only a linear combination of vector that still have $x_0$ as root
        modulo $N^m$.
    \item If we craft the lattice properly, the norm of shortest vector on the reduced basis will satisfy \ref{eq:pol}.
\end{itemize}

We can easily create polynomials $p_i$ ($g_{i,j}$ and $h_i$)  sharing the same root $x_0$ over $N^m$ of $f$ where $\delta$ is the degree of $f$:

\begin{center}
    \begin{eqnarray}
        g_{i,j}(x) &= x^j \cdot N^i \cdot f^{m-i}(x) \text{ for } i = 0,\hdots,m-1,\hspace{2mm} j=0,\hdots,\delta-1 \label{eq:create_pol} \\
        h_i(x) &= x^i \cdot f^m(x) \text{ for } i = 0,\hdots,t-1 
    \end{eqnarray}
\end{center}

@@

Suppose that we don't know the factorization of $N$ and we would like to know the message $m$.
However we know the padding and that the length of $m < 2^{16}$.

@@

which is the known padding string that got encrypted.

@@

\begin{align*}
    f(x) = x^3 + 0x15d393c596142306bae000000x^2 + 0x969acdb25c5f7b5ff7f29bc59x \\
    + 0x75abf10edc37c4b90c5897322
\end{align*}

@@

The rows of the matrix correspond to the coefficient vectors of the polynomials $f(x), Nx^2, Nx$ and $N$, furthermore we know
that each polynomials will be 0 modulo $N$ if evaluated at $x = m$.
We didn't fully applied Howgrave-Graham because we only used $m=1$ (the $N^m$ parameter not the message).\\
With this lattice construction every vector is of the form $v = (v_1X^3, v_2X^2, v_1X, v_0)$, because any
integer linear combination of the vector of the lattice will keep the bound $X^i$ for $i=0,..,\dim(B)-1$.

\textbf{Apply LLL}. We then apply LLL to find the shortest vector of the reduced basis:

\[
    \begin{split}
    v = (0x24409ad85f781X^3, -0xd75ab858b980X^2, \\
        0x1973847f4ddb8X, -0x9a2d87b43d)
    \end{split}
\]

We can construct the polynomial $g$ using the coefficients of $v$

\[
    \begin{split}
        g(x) = 0x24409ad85f781x^3  - 0xd75ab858b980x^2 \\ 
        + 0x1973847f4ddb8x - 0x9a2d87b43d
    \end{split}
\]

We know that

\[
    g(x_0) \equiv 0 \mod N, |x_0| \le X
\]

What we need to prove is that

\[
    \lVert g(xX) \rVert \le \frac{N}{\sqrt{n+1}}
\]

In this example, det $B = X^6N^3$, and LLL will find a short vector with $\lVert v \rVert \le 1.02^n (\det B)^{\frac{1}{dimB}}$.
If we ignore the $1.02^n$ factor, then we need to satisfy that

\[
    g(m) \le \lVert v \rVert \le (\det B)^{\frac{1}{4}} < X
\]

We have $(\det B)^{\frac{1}{4}} = (X^6N^3)^{\frac{1}{4}} < N$, if we solves for $X$ this will be satisfied when $X < N^{\frac{1}{6}}$.
In this case we have $N = 100$-bit number and $X = 2^{16}$, so $N^{\frac{1}{6}} > X$ and we have verified the correctness.\\

If we compute the root of $g(x)$ over the integers we obtain $m = 0x61$ which is the correct result.\\

If you want to find bigger root you need to create a bigger lattice with more polynomials generated with \ref{eq:create_pol},
this method is better described in TODO.
